---
title: "Reproducible research - Assignment week 2"
author: "Amelie"
date: "23 aout 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```

## Loading and preprocessing the data
Download the data from the internet
```{r Download data}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, destfile = "activity.zip")
unzip("activity.zip")
```
Preprocess the data
```{r Preprocess the data}
activity <- read.csv("activity.csv")
activity$Date <- as.Date(activity$date, format="%Y-%m-%d")
activity$Date <- as.POSIXct(activity$Date)
```

## What is mean total number of steps taken per day?
```{r Load librairies, results="hide"}
library("dplyr")
library("lubridate")
```
Calculate the total number of steps taken a day
```{r Steps per day - Sum}
sum_steps_day <- activity %>% 
  group_by(year = year(Date), month = month(Date), day = day(Date)) %>% 
  summarise(sum = sum(steps))
```
Create the histogram
```{r Steps per day - Histogram}
 hist(sum_steps_day$sum, main = "Number of steps taken per day", xlab = "Steps")
```  

Calculate the mean total number of steps taken a day
```{r Steps per day - Mean}
mean_tot_steps_day <- mean(sum_steps_day$sum, na.rm = TRUE)
mean_tot_steps_day
```  
```{r Steps per day - Median}
median_tot_steps_day <- median(sum_steps_day$sum, na.rm = TRUE)
median_tot_steps_day
```  

## What is the average daily activity pattern?
Average number of steps taken for all days per 5-minute interval
```{r Steps per interval - Mean}
sum_steps_int <- activity %>% 
  group_by(interval = interval) %>% 
  summarise(mean = mean(steps, na.rm=TRUE))
```
Create the plot
```{r Steps per interval - Plot}
plot(sum_steps_int$interval, sum_steps_int$mean, type="l", ylab = "Mean", xlab = "Interval", main = "Average number of steps taken for all days per 5-minute interval")
```  

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r Steps per interval - Max interval}
sum_steps_int[sum_steps_int$mean==max(sum_steps_int$mean),1]
```
## Imputing missing values
Calculate the total number of missing values in the dataset
```{r NA - Sum}
sum(is.na(activity$steps))
```
Filling missing values in a new dataset
```{r NA - Filling}
activity_noNA <- read.csv("activity.csv")
activity_noNA$Date <- as.Date(activity_noNA$date, format="%Y-%m-%d")
activity_noNA$Date <- as.POSIXct(activity_noNA$Date)
for(i in 1:nrow(activity_noNA)){
if(is.na(activity_noNA[i,1])){
    int <- activity_noNA[i,3]
    fill <- sum_steps_int[sum_steps_int$interval==int,2]
    activity_noNA[i,1] = fill}}
```
Calculate the total number of steps taken a day including imputed values
```{r Steps per day - Sum without NA}
sum_steps_day_noNA <- activity_noNA %>% 
  group_by(year = year(Date), month = month(Date), day = day(Date)) %>% 
  summarise(sum = sum(steps))
```
Create the histogram including imputed values
```{r Steps per day - Histogram without NA}
 hist(sum_steps_day_noNA$sum, main = "Number of steps taken per day", xlab = "Steps")
```  

Mean of steps taken per day including imputed values
```{r Steps per day - Mean without NA}
mean_tot_steps_day_noNA <- mean(sum_steps_day_noNA$sum)
mean_tot_steps_day_noNA
```  
Median of steps taken per day including imputed values
```{r Steps per day - Median without NA}
median_tot_steps_day_noNA <- median(sum_steps_day_noNA$sum)
median_tot_steps_day_noNA
```  
What is the impact of imputing missing data?
On the mean: the difference of *data with na* - *data without na* is:
```{r NA - Mean difference}
mean_tot_steps_day-mean_tot_steps_day_noNA
```
On the media: the difference of *data with na* - *data without na* is:
```{r NA - Median difference}
median_tot_steps_day-median_tot_steps_day_noNA
```
## Are there differences in activity patterns between weekdays and weekends?
Add the factor variable
```{r Patterns - Factors}
Sys.setlocale("LC_TIME", "English")
weekend_list <- c('Saturday', 'Sunday')
activity_noNA$wd_we <- factor((weekdays(activity_noNA$Date) %in% weekend_list), 
         levels=c(TRUE, FALSE), labels=c('weekend', 'weekday')) 
```

```{r Steps per interval and wdwe- Mean}
sum_steps_int_wdwe <- activity_noNA %>% 
  group_by(interval = interval, wdwe=wd_we) %>% 
  summarise(mean = mean(steps))
```
Create the plot
```{r Patterns - Plot}
library("ggplot2")
qplot(interval,mean, 
      data = sum_steps_int_wdwe,
      facets = .~wdwe, 
      geom = "line",
      main = "Average number of steps taken",
      xlab = "Interval", 
      ylab = "Steps")
```

